Trade-offs between sustainable development goals in systems of cities

### 2021-12-20 paper.tex
submit_paper

\author[label1]{Juste Raimbault\corref{cor1}}
\author[label2]{Denise Pumain}

@@

@@

The model parameters left free for optimisation are
\begin{enumerate}
    \item spatial interaction range for innovation diffusion
    \item spatial interaction range for population growth
    \item mutation (innovation) rate
    \item level of hierarchy to select cities inventing a new innovation (scaling of innovations with regard to urban size)
    \item rate of early adopters
    \item standard-deviation of new innovation utilities
    \item type of distribution for new innovation utilities.
\end{enumerate}

@@

We consider the ``innovation'' SDG (goal 9) and the ``climate'' SDG (goal 14) as conflicting objectives. We can expect that a higher economic activity linked to more intensive innovative activities will increase endogenous emissions, but also transport emissions between urban areas, generated by economic and transport flows. The existence of a trade-off remains an hypothesis which will be checked during the optimisation stage (here ``optimisation'' means trying to minimise simultaneously different stylised output indicators of the model).

\cite{chen2020does}

We consider therefore the two following objectives for model optimisation: (i) aggregated total utility during model dynamics, computed over time and across cities, with shares of each innovations; (ii) total emissions due to transport flows between cities, computed as cumulative population gravity flows.
% index of mobility intensity
% emission" est un terme englobant toute production ou processus qui tendrait à aggraver  l'épuisement des ressources et le changement climatique liés à la croissance des villes

@@

We investigate trade-offs between total innovation utility and emissions, by optimising the model using a bi-objective heuristic with free parameters and indicators detailed above. We use a NSGA2 optimisation algorithm, provided by OpenMOLE, with a population of 100 individuals, for 10,000 generations.
% 3.2 préciser ce que sont les "générations

\begin{figure*}
	\centering
	\includegraphics[width=\linewidth]{figures/pareto-oppAverageUtility-averageGravityFlow_color-gravityDecay_size-innovationDecay_shape-utilityDistribution.png}
	\caption{Pareto front between the opposite of average utility and total gravity flows. Point color gives population spatial interaction range; point size innovation diffusion range; and point shape the utility distribution.\label{fig:fig1}}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=\linewidth]{figures/pareto-oppAverageUtility-averageGravityFlow_VARYINGHIERARCHY_color-gravityDecay_size-innovationDecay.png}
	\caption{Pareto fronts, with initial population hierarchy fixed at different values (point shape).\label{fig:fig2}}
\end{figure*}

%  mentionner un "hierarchy index" en le légendant comme slope of the rank size
We now turn to experiments which could potentially provide policy insights. We run the same optimisation as before, but changing the initial population hierarchy of cities. To put it simply, we investigate how trade-offs change in different hypothetical systems of cities, ranging from highly hierarchical (Zipf exponent of 1.5) to a more balanced system (exponent of 0.5). We show results in Fig.~\ref{fig:fig2}. We find that the higher the hierarchical inequalities, the less flat the front. Overall, the less hierarchical system dominates the others (but this comparison remains limited as total population is different across systems). Furthermore, the size of the front is the smallest with the less unequal hierarchy, meaning that this system is indeed closer to some global optimum.

>>>>

%\author{}
\author[label1]{Juste Raimbault\corref{cor1}}
\author[label2]{Denise Pumain}

@@

%\end{document}

@@

The model parameters left free for optimisation are

\begin{enumerate}
    \item spatial interaction range for innovation diffusion
    \item spatial interaction range for population growth
    \item mutation (innovation) rate
    \item level of hierarchy to select cities inventing a new innovation (scaling of innovations with regard to urban size)
    \item rate of early adopters
    \item standard-deviation of new innovation utilities
    \item type of distribution for new innovation utilities.
\end{enumerate}

Some parameters can be parametrised from real data in a rather straightforward way, such as spatial interaction ranges by fitting spatial interaction models on appropriate data for example \citep{fotheringham1989spatial}, or the level of hierarchy for new innovation by fitting urban scaling laws \citep{pumain2006evolutionary}. Other parameters such as innovation rate or the distribution of innovation utilities correspond to a more abstract formulation which can not directly be linked to real-world proxies (for example, innovations compete along a single dimension). Furthermore, some parameters can be linked to potential policies while others can difficultly be acted upon. We choose thus to work with most parameters free to maximise the degrees of freedom explored by the optimisation algorithms, in some sense explore a broader set of scenarios for urban systems.

@@

We consider the ``innovation'' SDG (goal 9) and the ``climate'' SDG (goal 14) as conflicting objectives. We can expect that a higher economic activity linked to more intensive innovative activities will increase endogenous emissions, but also transport emissions between urban areas, generated by economic and transport flows. Empirical evidence does not suggest globally a simultaneous reduction of emissions through innovation \citep{chen2020does}. A potential decoupling of economic activity and emissions remains also still difficult to observe \citep{haberl2020systematic}. For these reasons, we can expect a compromise between these two dimensions. The effective existence of a trade-off in synthetic urban dynamics generated by the model remains in that context an hypothesis, which will be checked during the optimisation stage (here ``optimisation'' means trying to minimise simultaneously different stylised output indicators of the model).

We consider therefore the two following objectives for model optimisation:

\begin{enumerate}
 \item aggregated total utility during model dynamics, computed over time and across cities, with shares of each innovations, as
 \[
 U = \sum_{t,i,c} \delta_{t,i,c} \cdot \frac{P_{t,i}}{P_t} \cdot u_c
 \]
 where $t$ are time steps, $i$ city index, $c$ innovation index, $\delta_{t,i,c}$ innovation adoption shares, $P_{t,i}$ city population, $P_t$ total population, and $u_c$ innovation utility
 \item total emissions due to transport flows between cities, computed as cumulative population gravity flows; this indicators can be understood as some index of ``mobility intensity'' and used as a proxy for emissions; it is computed as
 \[
 E = \sum_{t,i,j} \frac{P_{t,i}P_{t,j}}{P_t^2} \cdot \exp\left(- d_{ij} / d_G\right)
 \]
 where $d_{ij}$ is the geographical distance between cities $i,j$ and $d_G$ population growth spatial interaction range.
\end{enumerate}

Note that in this abstract model and with the proxy used, ``emissions'' captures any production or process contributing to GHG emissions or to resource exhaustion linked to urban growth. More detailed indicators parametrised with empirical data remain to be explored.

@@

We investigate trade-offs between total innovation utility and emissions, by optimising the model using a bi-objective heuristic with free parameters and indicators detailed above. We use a NSGA2 optimisation algorithm, provided by OpenMOLE, with a population of 100 individuals, for 10,000 generations. The genetic algorithm proceeds iteratively, by progressively selecting an optimal population of individuals (parameter points), constructed from the previous generation of optimal individuals. We observe convergence when the hypervolume of the Pareto front becomes steady, what we observed in practice with this total number of runs.

\begin{figure*}
	\centering
	\includegraphics[width=\linewidth]{figures/pareto-oppAverageUtility-averageGravityFlow_color-gravityDecay_size-innovationDecay_shape-utilityDistribution.png}
	\caption{Pareto front between the opposite of average utility (innovation) and mobility intensity index (emissions). Point color gives population spatial interaction range; point size innovation diffusion range; and point shape the utility distribution.\label{fig:fig1}}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=\linewidth]{figures/pareto-oppAverageUtility-averageGravityFlow_VARYINGHIERARCHY_color-gravityDecay_size-innovationDecay.png}
	\caption{Pareto fronts, with initial population hierarchy index fixed at different values (point shape).\label{fig:fig2}}
\end{figure*}

We now turn to experiments which could potentially provide policy insights. We run the same optimisation as before, but changing the initial population hierarchy of cities. To put it simply, we investigate how trade-offs change in different hypothetical systems of cities, ranging from highly hierarchical (Zipf exponent of 1.5) to a more balanced system (exponent of 0.5). We expect to obtain different Pareto fronts, for different values of a hierarchy index, which corresponds to the slope of the initial Zipf law.

We show results in Fig.~\ref{fig:fig2}. We find that the higher the hierarchical inequalities, the less flat the front. Overall, the less hierarchical system dominates the others (but this comparison remains limited as total population is different across systems). Furthermore, the size of the front is the smallest with the less unequal hierarchy, meaning that this system is indeed closer to some global optimum.
